---
title: "divent"
subtitle: "Measures of Diversity based on Entropy"
bibliography: ../inst/REFERENCES.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to divent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r global_options, include=FALSE}
set.seed(97310)
```

_divent_ is a package for _R_ designed to estimate diversity based on HCDT entropy or similarity-based entropy. 
It is a reboot of the _entropart_ package, following the tidyverse manifest and easier to use.
This is a short introduction to its use.

The package allows estimating biodiversity according to the framework based on HCDT entropy, the correction of its estimation-bias [@Grassberger1988; @Chao2003; @Chao2015] and its transformation into equivalent numbers of species [@Hill1973; @Jost2006; @Marcon2014a].
Estimation of diversity at arbitrary levels of sampling, requiring interpolation or extrapolation [@Chao2014] is available

Phylogenetic or functional diversity [@Marcon2014b] can be estimated, considering phyloentropy as the average species-neutral diversity over slices of a phylogenetic or functional tree [@Pavoine2009]. 

Similarity-based diversity [@Leinster2012] can be used to estimate [@Marcon2014e] functional diversity from a similarity or dissimilarity matrix between species without requiring building a dendrogram and thus preserving the topology of species  [@Pavoine2005a; @Podani2007].

The classical diversity estimators (Shannon and Simpson entropy) can be found in many R packages. 
Bias correction is rarely available except in the _EntropyEstimation_ [@Cao2014] package which provides the Zhang and Grabchak's estimators of entropy and diversity and their asymptotic variance (not included in _entropart_).


# Estimating the diversity of a community

## Community data

Community data is a numeric vector containing abundances of species (the number of individual of each species) or their probabilities (the proportion of individuals of each species, summing to 1).

Example data is provided in the dataset `paracou_6_abd`.
Let's get the abundances of tree species in the 6.25-ha tropical forest plot #6 from Paracou forest station in French Guiana.
It is divided into 4 equally-sized subplots:

```{r load_paracou6}
library("divent")
paracou_6_abd
abd_sum(paracou_6_abd)
```


The data in `paracou_6_abd` is an object of class `abundances`, i.e. a tibble with species as columns and sites as rows.
It can be manipulated as any dataframe and plotted as a rank-abundance curve:

```{r plot_paracou6}
autoplot(paracou_6_abd[1, ])
```

The `rcommunity` function allows drawing random communities, e.g. a log-normal one [@Preston1948]:

```{r rcommunity}
rc <- rcommunity(1, size = 10000, distribution = "lnorm")
autoplot(rc, fit_rac = TRUE, distribution="lnorm")
```

The Whittaker plot (rank-abundance curve) of a random log-normal distribution of 10000 individuals simulated with default parameter ($\sigma = 1$) is produced.


## Diversity estimation

The classical indices of diversity are richness (the number of species), Shannon's and Simpson's entropies:

```{r shannon}
div_richness(paracou_6_abd)
ent_shannon(paracou_6_abd)
ent_simpson(paracou_6_abd)
```

When applied to probabilities (created with `as_probaVector` in the following example), no estimation-bias correction is applied: this means that indices are just calculated by applying their definition function to the probabilities (that is the naive, or plugin estimator).

```{r naiveshannon}
library("dplyr")
paracou_6_abd %>% 
  as_probabilities() %>% 
  ent_shannon()
```

When abundances are available, many estimators can be used [@Marcon2015a] to address unobserved species and the non-linearity of the indices:

```{r IndicesAbd}
ent_shannon(paracou_6_abd)
ent_shannon(paracou_6_abd, estimator = "ChaoJost")
```

The best available estimator is chosen by default: its name is returned.

Those indices are special cases of the Tsallis entropy [-@Tsallis1988] or order $q$ (respectively $q=0,1,2$ for richness, Shannon, Simpson):

```{r Tsallis}
ent_tsallis(paracou_6_abd, q = 1)
```

Entropy should be converted to its effective number of species, i.e. the number of species with equal probabilities that would yield the observed entropy, called @Hill1973 numbers or simply diversity [@Jost2006].

```{r Diversity}
div_hill(paracou_6_abd, q = 1)
```

Diversity is the deformed exponential of order $q$ of entropy, and entropy is the deformed logarithm of of order $q$ of diversity:

```{r lnq}
(d2 <- div_hill(paracou_6_abd, q = 2)$diversity)
ln_q(d2, q = 2)
(e2 <-ent_tsallis(paracou_6_abd, q = 2)$entropy)
exp_q(e2, q = 2)
```

Diversity can be plotted against its order to provide a diversity profile. Order 0 corresponds to richness, 1 to Shannon's and 2 to Simpson's diversities:

```{r DiversityProfile}
div_profile(paracou_6_abd) %>% autoplot
```


# Full documentation

https://ericmarcon.github.io/entropart/ 

# References
