---
title: "Phylogenies in divent"
output:
  rmarkdown::html_document:
    toc: yes
    toc_float: yes
bibliography: ../../inst/REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{Phylogenies in entropart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r global_options, include=FALSE}
set.seed(97310)
```

_divent_ is a package for _R_ designed to estimate diversity based on HCDT entropy or similarity-based entropy. 

Several object classes can be used in the package to represent phylogenies and calculate phylogenetic diversity. 
They are detailed here.


# Formats used

Supported phylogenies are classes `phylo` from package _ape_, `phylog` from package _ade4_ and `hclust` from package _base_.

To optimize performance, a specific class `phylo_divent` is introduced by the package: it is basically a list containing a `phylo` and an `hclust` representation of the phylogeny and preprocessed useful statistics.

The argument `tree` used in phylogenetic diversity estimation functions of the package may be any of those formats.
Conversion to `phylo_divent` is managed internally.


# Conversion between classes

Let's start from an ultrametric distance matrix.
Three species are defined, with distance 1 between the first 2 and distance 2 between both and the last one.

```{r DistMatrix}
species_dist <- matrix(c(0, 1, 2, 1, 0, 2, 2, 2, 0), nrow=3, byrow=TRUE)
row.names(species_dist) <- colnames(species_dist) <- c("A", "B", "C")
species_dist
```

## hclust

An `hclust` object is created by UPGMA hierarchical clustering.

```{r hclust}
library("stats")
plot(
  tree.hclust <- hclust(
    as.dist(species_dist), 
    method = "average"
  ), 
  hang = -0.01, 
  axes = F
)
axis(2)
```

Node heights are stored in `$height`.

```{r hclust2}
tree.hclust$height
```

## phylo

Conversion to a `phylo` object is straightforward.

```{r phylo}
library("ape")
plot(tree.phylo <- as.phylo(tree.hclust))
axis(1)
```

Edge lengths have been divided by 2 during the conversion, as documented in `?as.phylo.hclust`.

```{r phylo2}
tree.phylo$edge.length
```

That does not suit our needs.
In _divent_, edge lengths are multiplied by 2 after conversion, so that `phylo` objects can be identical to other phylogenies. 

```{r phylo3}
tree.phylo$edge.length <- 2 * tree.phylo$edge.length
plot(tree.phylo)
axis(1)
```

## phylog

The last conversion is from `phylo` to `phylog`.

```{r phylog}
library("ade4")
plot(tree.phylog <- hclust2phylog(tree.hclust))
axis(1)
```

Edge lengths are not stored directly in `phylog` objects.
The `$droot` component keeps distances from tips to root.
The `$Wdist` component is a `dist` (half a distance matrix without the zero-diagonal) whose values are $d_{s,t}=\sqrt{2 \delta_{s,t}^2}$, where $\delta_{s,t}$ is the distance between species $s$ ant $t$.

```{r phylog2}
tree.phylog$droot
tree.phylog$Wdist^2/2
```

`phylog` trees are supported by _divent_ but are converted to `hclust`:

```{r}
tree2.hclust <- stats::hclust(tree.phylog$Wdist^2 / 2, "average")
plot(tree2.hclust, hang = -0.01, axes = F)
axis(2)
```


## phylo_divent

The function `as_phylo_divent` converts any object of one of the previous types.

```{r phylo_divent}
library("divent")
plot(tree.phylo_divent <- as_phylo_divent(tree.phylo))
axis(2)
```

Its `plot` method is that of `dendrogram` objects.

A `phylo_divent` tree is a list.

```{r phylo_divent2}
tree.phylo_divent
```

It contains: 

- `$phylo`: the tree as a `phylo` object.
- `$hclust`: the tree as a `hclust` object.
- `$height`: the height of the tree.
- `$cuts`: the heights of the nodes of the tree.
- `$intervals`: the intervals between the cuts.
- `$phylo_groups`: a matrix that describes the ancestors of species in each interval.
The column names give the most recent cut of each interval.
Values are the group each present species belongs to in each interval.
In the example above, species A and B belong to the same group in the interval starting at 1. 


# Newick dendrograms

A Newick description of a dendrogram can be read into a `phylo` object, that can be in turn converted to `phylo_divent` or used directly in _divent_ functions.

```{r Newick}
NewickABC <- "(C:2,(A:1,B:1):1);"
plot(phyloABC <- ape::read.tree(text = NewickABC))
axis(1)
```
