---
title: "Random communities"
output:
  rmarkdown::html_document:
    toc: yes
    toc_float: yes
bibliography: ../../inst/REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{Random communities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
set.seed(97310)
# Installation of packages if necessary
install_packages <- function(packages) {
  install_package <- function(package) {
    if (!package %in% installed.packages()[, 1]) {
      install.packages(package, repos = "https://cran.rstudio.com/")
    }
  }
  invisible(sapply(packages, install_package))
}
install_packages(c("spatstat"))
```

The package proposes several functions to simulate random communities.


# Simulating abundances according to classical models

The classical models of ecology are implemented in **divent** in the `rcommunity()`.

To write:

- models


# Simulating abundances according to an observed abundance distribution

To write:

- multinomial 
- unveiled


# Simulating spatialized communities

## Communities

To write.

## Species

To simulate communities more accurately, species can be drawn one by one.
This way, spatial structures and size distributions can be chosen for each species.

The following example shows how to do it: a 1-ha forest made of an aggregated species of interest and undifferentiated other ones.
`rspcommunity()` is called with arguments `n = 1` and `species_number = 1`.
A list is of point patterns is obtained and superimposed to make a single `wmppp` object.

```{r}
library("spatstat")
library("dbmss")
library("ggplot2")
library("dplyr")
library("divent")
# A 1-ha window
the_win <- square(100, unitname = c("meter", "meters"))
# Simulate species
the_species <- do.call(
  superimpose,
  list(
    rspcommunity(
      # A "grey" species, i.e. all individuals of species out of focus
      n = 1, species_number = 1, species_names = "others", size = 500,
      # Spatial features: binomial distribution
      win = the_win, spatial = "Binomial",
      # Size distribution
      weight_distribution = "Exponential", w_min = 10, w_mean = 15),
    rspcommunity(
      # The species of interest
      n = 1, species_number = 1, species_names = "aggregated", size = 50,
      # Spatial features: aggregated distribution
      win = the_win, spatial = "Thomas", thomas_scale = 10, thomas_mu = 20,
      # Size distribution
      weight_distribution = "Weibull", weibull_scale = 30, weibull_shape = 2,
      w_min = 10
    )
  )
)
# Map the simulated community
autoplot(the_species) +
  labs(color = "Species", size = "Diameter")
```

Post-simulation processing of such spatialized communities may be for instance thinning to reduced the local average basal area.

```{r}
# Basal area
G <- (spatstat.geom::marks(the_species)$PointWeight)^2 * pi / 4E4
sum(G)
# Density
G_density <- density(the_species, sigma = 10, weights = G)
plot(G_density)

# Thinning: the target is at most 40m² locally
target <- mean(G_density) * 40 / sum(G)
thin_im <- G_density
thin_im$v <- pmin(pmax(target / G_density$v, 0), 1)
# Plot the thinning probability
plot(thin_im)
# Plot the thinned community 
rthin(
  the_species,
  P = thin_im
) %>% autoplot()
```

