---
title: "Random communities"
output:
  rmarkdown::html_document:
    toc: yes
    toc_float: yes
bibliography: ../../inst/REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{Random communities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
set.seed(97310)
# Installation of packages if necessary
install_packages <- function(packages) {
  install_package <- function(package) {
    if (!package %in% installed.packages()[, 1]) {
      install.packages(package, repos = "https://cran.rstudio.com/")
    }
  }
  invisible(sapply(packages, install_package))
}
install_packages(c("spatstat"))
```

The package proposes several functions to simulate random communities.


# Simulating abundances according to classical models

The classical models of ecology are implemented in **divent** in the `rcommunity()` function.

Arguments are :

- `n`: the number of communities to draw.
- `size`: the number of individuals to draw in each community.
- `species_number`: the number of species.
- `distribution`: the distribution of species abundances.

Distributions may be "lnorm" (log-normal), "lseries" (log-series), "geom" (geometric) or "bstick" (broken stick):

- log-normal ("lnorm") distributions [@Preston1948] require argument `sd_lnorm` to set the standard deviation of the log-abundances. 
The expectation of the distribution is computed from the other arguments.
- log-series ("lseries") distributions [@Fisher1943] rely on the `fisher_alpha` argument.
They ignore `species_number` because the number of species is given by  `fisher_alpha` and `size`.
-  geometric ("geom") distributions [@Motomura1932] have argument `prob_geom`.
This distribution is not stochastic.
- MacArthur's broken stick ["bstick", @MacArthur1957] has no argument.

The probability distribution is first drawn, once.
Abundances are then drawn in `n` multinomial distributions, with respect to the probabilities and `size`.

Example: a single log-normal community with 300 species and standard deviation equal to 2 is drawn and plotted:

```{r}
library("dplyr")  # for `%>%`
library("divent")
rcommunity(1, size = 1E4, species_number = 300, distribution = "lnorm") %>% 
  # Rank-abundance plot
  autoplot()
```


# Bootstrapping an observed abundance distribution

Bootstrapping allows drawing many community according to the same probability distribution according to a multinomial law with respect to the probabilities and `size`.
This is useful to compute confidence intervals of any statistics of the distribution.

Arguments are :

- `n`: the number of communities to draw.
- `size`: the number of individuals to draw in each community.
- `prob` or `abd`: a numeric vector containing probabilities or abundances.
One of them must be given, the other one must be `NULL`.
- `bootstrap`: defines the technique used:
    - "Marcon2012": following @Marcon2012a, abundances are just normalized to obtain probabilities.
    If `prob` is given rather than `abd`, these probabilities are used.
    - "Chao2013" or "Chao2015" (by default): based on the observed abundances (`abd` must be given), the actual probability distribution is estimated following @Chao2013 or @Chao2015.
    Observed frequencies (normalized abundances) are not directly used as estimators of the probabilities: abundant species are correctly estimated but rare species are either overestimated or unobserved.
    A model based on sample coverage is used to correct this issue.
    The estimator of sample coverage is chosen by the argument `coverage_estimator`: the sum of observed probabilities equals the sample coverage.
    The number of unobserved species is estimated and their distribution (their probabilities sum to the coverage deficit) is arbitrary set as uniform ("Chao2013") or geometric ("Chao2015").

Example: a community based on the subplot 1 of Paracou plot 6 is drawn and its distribution is compared to the observed frequencies.

```{r}
library("ggplot2") # for `labs()`
# Draw a large community to avoid sample bias
rcommunity(1, size = 1E6, abd = as.numeric(paracou_6_abd[1, ])) %>% 
  # Compute the probabilities
  as_probabilities() %>% 
  # Rename the distribution
  mutate(site = "bootstrap") %>% 
  # Compare to the original distribution
  bind_rows(as_probabilities(paracou_6_abd[1, ])) %>% 
  autoplot() +
  labs(color = "Distribution")
```

The probabilities of abundant species are not modified.
The probabilities of rare species is smaller than their frequencies and more than a hundred species are added.


# Simulating spatialized communities

## Communities

Spatialized communities include the location of individuals in a window.
The abundance distribution is drawn according to the same arguments as in `rcommunity()`.

Additional arguments allow choosing:

- the window: `win` must be an `owin` object from the **spatstat.geom** package.
Default is a square window whose side length is 1.
- the species names: `species_names` is a character vector with the species names.
They may be generated automatically.
- the spatial distribution of the points in argument `spatial`, i.e. the point process used to draw their location:
    - "Binomial" draws points uniformly in the window.
    - "Thomas" draws aggregated points: `thomas_mu` is the average number of points per cluster and `thomas_scale`  the standard deviation of random displacement (along each coordinate axis) of a point from its cluster center.
- the `weight_distribution` of the points sets their sizes:
    - "Uniform": a uniform distribution between `w_min` and `w_max`.
    This is the default distribution, with `w_min = w_max = 1.`
    - "Weibull": parameters are `w_min` (minimum size), `weibull_scale` (the scale parameter) and shape (the shape parameter).
    - "Exponential": argument `w_mean` is the mean weight, i.e. the negative of the inverse of the decay rate.

Example: a random community is drawn and mapped.
It contains 100 trees of 5 species. 
Their spatial distribution is aggregated and their size distribution is exponential (with default parameters).

```{r}
rspcommunity(
  1, 
  size = 100, 
  species_number = 5, 
  spatial = "Thomas", 
  weight_distribution = "Exponential"
) %>% 
  autoplot()
```

## Species

To simulate communities more accurately, species can be drawn one by one.
This way, spatial structures and size distributions can be chosen for each species.

The following example shows how to do it: a 1-ha forest made of an aggregated species of interest and undifferentiated other ones.
`rspcommunity()` is called with arguments `n = 1` and `species_number = 1`.
A list is of point patterns is obtained and superimposed to make a single `wmppp` object.

```{r}
library("spatstat") # for `square()`
library("dbmss")    # for `superimpose.wmppp()`
library("ggplot2")  # for `labs()`
library("dplyr")    # for `%>%`
library("divent")
# A 1-ha window
the_win <- square(100, unitname = c("meter", "meters"))
# Simulate species
the_species <- do.call(
  superimpose,
  list(
    rspcommunity(
      # A "grey" species, i.e. all individuals of species out of focus
      n = 1, species_number = 1, species_names = "others", size = 500,
      # Spatial features: binomial distribution
      win = the_win, spatial = "Binomial",
      # Size distribution
      weight_distribution = "Exponential", w_min = 10, w_mean = 15),
    rspcommunity(
      # The species of interest
      n = 1, species_number = 1, species_names = "aggregated", size = 50,
      # Spatial features: aggregated distribution
      win = the_win, spatial = "Thomas", thomas_scale = 10, thomas_mu = 20,
      # Size distribution
      weight_distribution = "Weibull", weibull_scale = 30, weibull_shape = 2,
      w_min = 10
    )
  )
)
# Map the simulated community
autoplot(the_species) +
  labs(color = "Species", size = "Diameter")
```

Post-simulation processing of such spatialized communities may be for instance thinning to reduced the local average basal area.

```{r}
# Basal area
G <- (spatstat.geom::marks(the_species)$PointWeight)^2 * pi / 4E4
sum(G)
# Density
G_density <- density(the_species, sigma = 10, weights = G)
plot(G_density)

# Thinning: the target is at most 40m² locally
target <- mean(G_density) * 40 / sum(G)
thin_im <- G_density
thin_im$v <- pmin(pmax(target / G_density$v, 0), 1)
# Plot the thinning probability
plot(thin_im)
# Plot the thinned community 
rthin(
  the_species,
  P = thin_im
) %>% autoplot()
```

